Gaia's constraints on the very massive (initially $\sim$$150\,M_\odot$) runaway star candidate VFTS682

### 2018-06-25 VFTS682_APJL.tex
selma

\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
  R.~P.~van~der~Marel$^{3,4}$, E.~L.~Laplace$^{1}$,
  J.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
  V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{1,8,9}$,  A.~de~Koter$^{1}$,
  N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~Schneider$^{13}$, J.~S.~Vink$^{14}$}
\affil{ $^{*}$ Corresponding author: \href{mailto:M.Renzo@uva.nl}{M.Renzo@UvA.nl}.  A list of the affiliations can be found at the end of this paper.}

@@

 How very massive stars form is still an intriguing open question in stellar astrophysics.  VFTS682 is among the most massive stars known, with an inferred initial mass of about  $150\,M_\odot$. It is located in 30 Doradus at a projected distance of 29\,pc from the central cluster R136.  The absence of other massive stars in its immediate vicinity led to two intriguing hypotheses posed in earlier work: either (a) it formed in relative isolation through a new mode of star formation or (b) it was ejected dynamically from the central cluster.

Aiming to shine light on this debate, we investigate the kinematics of
VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space
  Telescope} photometry. We derive a projected velocity relative of to
the cluster of $\sim$$30\pm20\kms$.  The direction and magnitude of
the proper motion are consistent with what is expected if VFTS682 was ejected as a slow runaway from the central cluster, given its distance and the inferred age.  However, the error bars are substantial and the proper motion measurements alone cannot rule out the counter hypothesis with high confidence.

If future data confirm the runaway nature, this would make the VFTS682
the most massive runaway star known to date.  While we cannot prove
this solidly from the current data alone, we do consider this
hypothesis as the most plausible because of a variety of
circumstantial clues. The central cluster is known to harbor several
other stars more massive than $150\,M_\odot$ similar in spectral type.
The very massive O stars VFTS16 and VFTS72, which have recently been identified as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
\end{abstract}

@@

In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to reside in the innermost cores. In this picture, field stars are primarily the result of the dissolution of dense groups.  However, a significant population of massive stars exists in relative isolation, far from dense clusters or OB associations and their origin remains matter of debate \citep{gvaramadze:12, lamb:16,ward:18}.   One hypothesis to explain the population of relatively isolated massive stars is that they formed in the field. The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.  Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, renzo:18}.

@@

The WNh5 star VFTS682, located at right ascension (RA)
05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
%
%of spectral type WNh5 using multi-epoch spectra covering $\lambda$4000--7000 from the
%VFTS survey. % The spectra available covered $\lambda$4000--7000 and
% were taken at multiple epochs.
%

\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
the brightest stars in the region. The star is unlikely to have a
close companion, because of the absence of clear radial velocity
variations \citet{bestenlehner:11}. Bayesian fits of the stellar
parameters against evolutionary tracks \citep{brott:11, kohler:15}
using the Bonsaii code \citep{schneider:17, schneider:18} provide
estimates for the age, present mass and initial mass, indicating that
this star among the most massive stars known, see
\Tabref{tab:star_param} for an overview of the parameters.

\cite{townsley:06} did not find
an X-ray point source at the position of VFTS682 in their
\emph{Chandra} survey of the region, which suggests the absence of
colliding winds that would be expected in the presence of companions
even for extreme mass ratios, given the extremely large mass of VFTS682.

@@

%of $137.8^{+27.5}_ {-15.9}\Msun$, an apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass of $150.0^{+28.7}_{-17.4}\Msun$.  Further parameters of are summarized in Table~\ref{X}.
VFTS682 is relatively isolated in \emph{Spitzer} images, except for the extended
source S10 interpreted as star formation triggered by the wind of
VFTS682 itself by \cite{walborn:13}.

>>>>

\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, \SdM{I.~Platais$^{3}$,}
  R.~P.~van~der~Marel$^{3,4}$,  \SdM{E.~L.~Laplace$^{1}$,}
  J.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
  V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{1,8,9}$,  A.~de~Koter$^{1}$,
  N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~Schneider$^{13}$, J.~S.~Vink$^{14}$}
\affil{ $^{*}$ Corresponding author: \href{mailto:M.Renzo@uva.nl}{M.Renzo@UvA.nl}.  A list of the affiliations can be found at the end of this paper.}

@@

 How very massive stars form is still among the most intriguing open  questions in stellar astrophysics.  VFTS 682 is among the most  extreme massive stars known, with an inferred initial mass of about  $150\,M_\odot$. It is located in 30 Doradus at a projected distance of 29\,pc from the central cluster R136.  The absence of other massive stars in its immediate vicinity led to two intriguing hypotheses posed in earlier work: either (a) it formed in relative isolation through a new mode of star formation or (b) it was ejected dynamically from the central cluster.

Aiming to shine light on this debate, we investigate the kinematics of  VFTS 682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space Telescope} photometry.    We derive a projected velocity relative of to the cluster of $\sim$$30\pm20\kms$.  The direction and magnitude of the proper is consistent with what is expected if VFTS 682 was ejected as a slow runaway from the central cluster, given its distance and the inferred age.  However, the error bars are substantial and the proper motion measurements alone cannot rule out the counter hypothesis with high confidence.

If future data confirms the runaway nature, this would make the VFTS682 the most massive runaway star known to date.  While we cannot prove this solidly from the current data alone, we do consider this hypothesis as the most plausible because of a variety of circumstantial clues. The central cluster is known to harbor several other stars more massive than $150\,M_\odot$ similar in spectral type.  The very massive O star VFTS 16, which has recently been identified as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions by numerical simulations.
\end{abstract}

@@

In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to reside in the innermost cores. In this picture, field stars are primarily the result of the dissolution of dense groups.  However, a significant population of massive stars exists in relative isolation, far from dense clusters or OB associations and their origin remains matter of debate \citep{gvaramadze:12, lamb:16,ward:18}.   One hypothesis to explain the population of relatively isolated massive stars is that they formed in the field. The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.  Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the \SdM{ companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, eldridge:11, renzo:18}}.

@@

The WNh5 star VFTS682, located at right ascension (RA)
05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}.

%
%of spectral type WNh5 using multi-epoch spectra covering $\lambda$4000--7000 from the
%VFTS survey. % The spectra available covered $\lambda$4000--7000 and
% were taken at multiple epochs.
%

\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
the brightest stars in the region. 
%
\SdM{The star does not show clear radial velocity variations, which makes the presence of a close companion unlikely
 \citet{bestenlehner:11}.  There are also no indications for the presence of a companion based on observations with the Chandra X-Ray Observatory. 
The detection of a point source in X-rays could have pointed at the presence of a compact companion or a colliding wind binary \citep{townsley:06}. }
% 
%\emph{Chandra} survey of the region, which suggests the absence of colliding winds that would be expected in the presence of companions even for extreme mass ratios, given the extremely large mass of VFTS682.  - Selma: that depends on the separation.  I don't know what the predictions are for very unequaly winds and the LMC is far away.  I shortened it a bit and moved this up to merge it where we were discussing the presence of a companion already, to use it in the flow of the argument, instead of disrupting the flow. 
%
 Bayesian fits of the stellar parameters derived from the spectra against single star evolutionary tracks \citep{brott:11, kohler:15}
using the Bonsaii code \citep{schneider:17, schneider:18} provide
estimates for the age, present mass and initial mass, indicating that
this star among the most massive stars known, see
\Tabref{tab:star_param} for an overview of the parameters.

@@

\SdM{VFTS682 also appears to be relatively isolated in (near) infrared.   There nearest bright (near) infrared sources detected by \emph{Spitzer}  \citep{meixner:06} and resolved in the \emph{VISTA} Magellanic Clouds Survey \citep{cioni:11} are located at a distance of about 10 arcsecond, i.e. about 2.4 pc. \cite{walborn:13} speculate that these nearby young stars may represent a case of starformation triggered by the the wind of VFTS682, but we chance alignment cannot be excluded.  }

% interpreted as star formation triggered by the wind of  VFTS682 itself by \cite{walborn:13}.   - SdM: too strong lets be more careful.  This is a controversial topic.  Walborn formulates more careful than you summarize him.

%of $137.8^{+27.5}_ {-15.9}\Msun$, an apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass of $150.0^{+28.7}_{-17.4}\Msun$.  Further parameters of are summarized in Table~\ref{X}.
